---
title: "Imputation evaluation for compositional geothermal fluid samples"
author: 'G. Santamaria-Bonfil'
output: 
  rmarkdown::github_document
  # html_document:
  #   toc: true
  #   toc_depth: 3
  #   toc_float: true
  #   number_sections: true
  #   df_print: paged
  #   keep_md: true
    # css: journal.css
---

# Libraries

```{r}
library(TOSTER)
library(stringr)
library(dplyr)
library(reshape)

library(ggplot2)
library(ggridges)
library(viridis)
library(scales)
library(latex2exp)
```

# Function to transform from PPM to Milliequivalents

```{r}
raw.2.mlEquivalent <- function(tmp_df)
{
  res_df <- tmp_df
  
  res_df$Na   <- tmp_df$Na/22.989770
  res_df$K    <- tmp_df$K/39.0983
  res_df$Mg   <- (tmp_df$Mg*2)/24.3050
  res_df$Ca   <- (tmp_df$Ca*2)/40.078
  res_df$Li   <- tmp_df$Li/6.941
  res_df$Cl   <- tmp_df$Cl/35.4527
  res_df$SO4  <- (tmp_df$SO4*2)/(32.066+(15.9994*4))
  res_df$HCO3 <- (tmp_df$HCO3)/(1.00794+12.0107+(15.9994*3))
  
  return(res_df)
}
```

# Transform from PPM to MilliEquivalents

The original and imputed WCGDb must be in the working directory.
But first, we require to locate the original and the imputed datasets.
For illustration purposes, the working directory is set to

"C:/Users/guillermo.santamaria/Documents/Results"

Modify this path to your convenience.

```{r setup}
knitr::opts_knit$set(root.dir = 'C:/Users/guillermo.santamaria/Documents/Results')
```

## Original Data

```{r}
orig_csv_name = 'original/simplifiedWCGDb.csv'
original.data <- read.csv(orig_csv_name, header = T, stringsAsFactors = F)

drop_col <- c('X')#Id column

tmp_ixes <- !(colnames(original.data) %in% drop_col)
ft_names <- colnames(original.data)[tmp_ixes]

original.data <- original.data[,ft_names]
original.data <- raw.2.mlEquivalent(original.data)
```

## Imputed Data

```{r}
imp_path<- paste0(getwd(),'/GFD Imputed DB/')

files_ls <- list.files(imp_path)
imp_dfs <- list()

for(i in 1:length(files_ls))
{
  # print(files_ls[i])
  imp_dfs[[i]] <- read.csv(paste0(imp_path,files_ls[i]), header = T)
  imp_dfs[[i]]['Temperature'] <- original.data$Temperature
  imp_dfs[[i]] <- imp_dfs[[i]][,ft_names]
  imp_dfs[[i]] <- raw.2.mlEquivalent(imp_dfs[[i]])
}
```

### Renaming imputed datasets by the name of the algorithm

```{r}
imp_file_names <- strsplit(files_ls, '_')
imp_names <- sapply(1:length(imp_file_names), 
                    function(i) imp_file_names[[i]][2])
names(imp_dfs) <- imp_names
```


```{r}
na_col_ixes <- sapply(1:dim(original.data)[2], 
                      function(i) any(is.na(original.data[,i])))

ckc4eqv.ft <- colnames(original.data)[na_col_ixes]

EqvlntImp.df <- data.frame(expand.grid(imp_names,ckc4eqv.ft))
colnames(EqvlntImp.df) <- c('algorithm', 'imputed.var')

EqvlntImp.df <- EqvlntImp.df %>% 
                mutate(algorithm.parsed = recode_factor(algorithm, 
                                                        'dt514a' = "DT^-", 
                                                        'dt514b' = "DT^*", 
                                                        'dt522' = "miDT^*", 
                                                        'knn' = "KNN^-",
                                                        'mean' = "mu",
                                                        'median' = "median",
                                                        'mice' = "MICE^-",
                                                        'rf522' = "RF^*",
                                                        'stocLR' = "StochLR",
                                                        'svrl514a' = "SVR_l^-", 
                                                        'svrl514b' = "SVR_l^*", 
                                                        'svrl522' = "miSVR_l^*",
                                                        'svrr514a' = "SVR_r^-", 
                                                        'svrr514b' = "SVR_r^*", 
                                                        'svrr522' = "miSVR_r^*"
                                                        ))

EqvlntImp.df$algorithm <- as.character(EqvlntImp.df$algorithm)
EqvlntImp.df$imputed.var <- as.character(EqvlntImp.df$imputed.var)

EqvlntImp.df$min.Eq.val <- 0.99
EqvlntImp.df$p.ttest    <- 0.001
EqvlntImp.df$p.tost     <- 0.99
EqvlntImp.df$has.eqvlnc <- F
```


```{r}
for(j in 1:dim(EqvlntImp.df)[1])
{
  tmp_df  <- imp_dfs[[EqvlntImp.df$algorithm[j]]]
  tmp_ft  <- EqvlntImp.df$imputed.var[j]
  ft_orig <- original.data[,tmp_ft]
  ft_imp  <- tmp_df[,tmp_ft]
  
  d <- seq(0.0001, 0.99, by = 0.025)
  df_tost <- data.frame()
  for (i in 1:length(d)) 
  {
    
    m1 <- mean(ft_orig,na.rm = T)
    m2 <- mean(ft_imp)
    
    sd1 <- sd(ft_orig,na.rm = T)
    sd2 <- ifelse(sd(ft_imp)==0,0.000000001,sd(ft_imp))
    
    n1 <- length(na.omit(ft_orig))
    n2 <- length(ft_imp)
    
    low_eqbound_d  <- -d[i]
    high_eqbound_d <- d[i]
    
    alpha <- 0.05
    
    tost.test <- TOSTtwo(m1, m2, 
                         sd1, sd2, 
                         n1, n2, 
                         low_eqbound_d, high_eqbound_d, 
                         alpha,
                         var.equal = F, plot = F, verbose = F)
    
    sdpooled <- sqrt((sd1^2 + sd2^2)/2)
    low_eqbound <- low_eqbound_d * sdpooled
    high_eqbound <- high_eqbound_d * sdpooled
    degree_f <- (sd1^2/n1 + sd2^2/n2)^2/(((sd1^2/n1)^2/(n1 - 
                                                          1)) + ((sd2^2/n2)^2/(n2 - 1)))
    t1 <- ((m1 - m2) - low_eqbound)/sqrt(sd1^2/n1 + sd2^2/n2)
    p1 <- pt(t1, degree_f, lower.tail = FALSE)
    t2 <- ((m1 - m2) - high_eqbound)/sqrt(sd1^2/n1 + sd2^2/n2)
    p2 <- pt(t2, degree_f, lower.tail = TRUE)
    t <- (m1 - m2)/sqrt(sd1^2/n1 + sd2^2/n2)
    pttest <- 2 * pt(-abs(t), df = degree_f)
    ptost <- max(p1, p2)
    
    df_tost <- rbind(df_tost, data.frame(tost.test))
    if(pttest > alpha && ptost <= alpha) 
    {
      EqvlntImp.df$min.Eq.val[j] <- d[i]
      EqvlntImp.df$p.ttest[j]    <- pttest
      EqvlntImp.df$p.tost[j]     <- ptost
      
      na_ix   <- is.na(ft_orig)
      
      #------------
      #Uncomment the following code if you want to observe the density function
      #of the original WCGDb vs the imputed dataset
      
      # plot(density(ft_orig[!na_ix]), col = "blue",
      #      main=EqvlntImp.df$algorithm[j], xlab =tmp_ft,
      #      cex.lab=1.4, cex.axis=1.4, cex.main=1.4, cex.sub=1.4 )
      # lines(density(ft_imp), col = "red")
      
      #------------
      
      EqvlntImp.df$has.eqvlnc[j] <- T

      break
    }
  }
}
```

## Minimun Equivalence Interval

```{r, fig.width=12, fig.height=7, fig.res = 300, echo=FALSE}
res <- EqvlntImp.df %>% #group_by(algorithm) %>%
                        group_by(imputed.var) %>% filter(min.Eq.val<=0.3)

print(res)

tmp1 <- EqvlntImp.df %>% 
                    group_by(algorithm) %>% 
                    mutate(No.Feats.Equivalent = length(algorithm[min.Eq.val<=0.3]))
tmp1$No.Feats.Equivalent <- as.factor(tmp1$No.Feats.Equivalent)

tmp.lim <- c('dt514a','dt514b','dt522','knn','mean','median','mice',
             'rf522','stocLR','svrl514a','svrl514b','svrl522',
             'svrr514a','svrr514b','svrr522')
tmp.lbs <- c('dt514a' = parse(text= TeX("$DT^{-}$")),
             'dt514b' = parse(text = TeX("$DT^{*}$")),
             "dt522" = parse(text = TeX("$miDT^{*}$")),
             'knn' = parse(text = TeX("$KNN^{-}$")),
             'mean' = parse(text = TeX("$\\mu$")), 
             "median" = parse(text = TeX("$median$")),
             "mice" = parse(text = TeX("$MICE^{-}$")), 
             'rf522' = parse(text = TeX("$RF^{*}$")),
             "stocLR" = parse(text = TeX("$SLR^{-}$")), 
             "svrl514a" = parse(text = TeX("$SVR_l^{-}$")), 
             "svrl514b" = parse(text = TeX("SVR_l^{*}")), 
             "svrl522" = parse(text = TeX("$miSVR_l^{*}$")), 
             "svrr514a" = parse(text = TeX("$SVR_r^{-}$")), 
             "svrr514b" = parse(text = TeX("SVR_r^{*}")), 
             "svrr522" = parse(text = TeX("$miSVR_r^{*}$")))

# png(paste0(savef.path,"mei_plot.png"), width=1800, height=1000, res = 160)
ggplot(tmp1, aes(x=reorder(algorithm, -min.Eq.val, FUN = mean), y=min.Eq.val, fill=No.Feats.Equivalent))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(shape = imputed.var), color="black", size=2, stroke = .9 )+
  scale_x_discrete(breaks=tmp.lim, labels = tmp.lbs)+
  scale_y_continuous(breaks = seq(0, 1,by=0.05))+
  coord_flip()+
  scale_fill_viridis(discrete = T, alpha=0.6, option="D")+
  theme(axis.line.x = element_line(size = 0.5, linetype = "solid",
                                   colour = "black"),
        axis.line.y = element_line(size = 0.5, linetype = "solid",
                                   colour = "black"),
        axis.text.x = element_text(size = 12,angle=45,hjust=1),
        axis.text.y = element_text(size = 13),
        axis.title.x = element_text(size = 13),
        axis.title.y = element_text(size = 13),
        legend.text = element_text(size = 11),
        legend.title = element_text(size = 12),
        panel.background = element_rect(fill='white', color='white'))+
  geom_hline(yintercept = 0.1, linetype="dashed", color = "black", size=0.5)+
  geom_hline(yintercept = 0.2, linetype="dashed", color = "black", size=0.5)+
  geom_hline(yintercept = 0.3, linetype="dashed", color = "black", size=0.5)+
  labs(fill = "No. of Equivalent Features", shape = 'Imputed Feature')+
  xlab('')+ylab('MEI')
```

### Ridgeline plots of original vs Imputed dataset
In order to appreciate the imputation quality carried out by the different 
imputation algorithms, the density distribution function of the original WCGDb 
against the datasets obtained by the imputation algorithms is shown.

```{r, echo=FALSE}
for(i in ckc4eqv.ft)
{
  ridge_df  <- data.frame(log(original.data[,i]))
  colnames(ridge_df) <- 'Original'
  achiev_eq <- EqvlntImp.df %>% filter(imputed.var==i & has.eqvlnc == T)
  achiev_eq <- achiev_eq[order(achiev_eq$min.Eq.val, decreasing = F),]
  
  for(k in 1:dim(achiev_eq)[1])
  {
    ridge_df <- cbind(ridge_df,
                      data.frame(log(imp_dfs[[achiev_eq$algorithm[k]]][,i])))
  }
  
  colnames(ridge_df)[2:dim(ridge_df)[2]] <- achiev_eq$algorithm
  
  ridge_mtl <- melt(ridge_df)
  
  plot(
  ggplot(ridge_mtl, aes(x = value, y = variable, fill = variable)) + 
    geom_density_ridges(scale = 2)+
    # scale_y_discrete(expand = c(0.01, 0)) +
    scale_x_continuous(expand = c(0.01, 0)) +
    scale_fill_viridis(name = 'Imputation Alg.', discrete=TRUE, option = 'C',direction = -1)+
    theme(legend.position = "none", panel.background = element_blank(), 
          text = element_text(size=10)) + labs(x="", y="" )+
    labs(title = i)
  )

}
```

# Charge Balance Error Analysis

```{r,fig.width = 12, fig.height = 7, fig.res = 300, echo=FALSE}
#
cbe_df <- data.frame(matrix(nrow=dim(original.data)[1],ncol=(length(imp_names)+1)))
colnames(cbe_df) <- c('original',imp_names)

cation_list <- c('Li', 'Ca', 'Mg', 'K', 'Na')
anion_list <- c('Cl', 'SO4', 'HCO3')

for(j in colnames(cbe_df))
{
  if(j=='original')
  {
    cation_sum <- rowSums(original.data[cation_list], na.rm=T)
    anion_sum  <- rowSums(original.data[anion_list], na.rm=T)
  }else{
    tempsy <- imp_dfs[[j]]
    cation_sum <- rowSums(tempsy[cation_list], na.rm=T)
    anion_sum  <- rowSums(tempsy[anion_list], na.rm=T)
  }
  CBE <- abs(((cation_sum-anion_sum)/(cation_sum+anion_sum))*100)
  cbe_df[,j] <- CBE
}

cbe_lim5 <- 5
cbe_lim10<- 10
cbe_lim20<- 20

ddata <- data.frame(cbind(cbe_lim5,cbe_lim10,cbe_lim20))
ddata.m <- melt(ddata)
colnames(ddata.m)[1] <- 'upper.lim'
ddata.m$upper.lim <- c('CBE \u2264  5', 'CBE \u2264 10', 'CBE \u2264 20')

cbe_mtl <- melt(cbe_df)
# cbe_metl1 <- cbe_mtl
cbe_metl1 <- filter(cbe_mtl, variable =='original' | variable =='knn' | variable =='dt514b' | variable =='svrr522' | variable == 'svrr514b' | variable == 'mice' | variable == 'mean')

tmp.lbs <- c('original'= 'Original', 
             'dt514a' = parse(text= TeX("$DT^{-}$")),
             'dt514b' = parse(text = TeX("$DT^{*}$")),
             "dt522" = parse(text = TeX("$miDT^{*}$")),
             'knn' = parse(text = TeX("$KNN^{-}$")),
             'mean' = parse(text = TeX("$\\mu$")), 
             "median" = parse(text = TeX("$median$")),
             "mice" = parse(text = TeX("$MICE^{-}$")), 
             'rf522' = parse(text = TeX("$RF^{*}$")),
             "stocLR" = parse(text = TeX("$SLR^{-}$")), 
             "svrl514a" = parse(text = TeX("$SVR_l^{-}$")), 
             "svrl514b" = parse(text = TeX("SVR_l^{*}")), 
             "svrl522" = parse(text = TeX("$miSVR_l^{*}$")), 
             "svrr514a" = parse(text = TeX("$SVR_r^{-}$")), 
             "svrr514b" = parse(text = TeX("SVR_r^{*}")), 
             "svrr522" = parse(text = TeX("$miSVR_r^{*}$")))

plot(
  ggplot(cbe_metl1, aes(value, color = variable, fill = variable)) +
  stat_ecdf(geom = "area",alpha = 0.15, color = NA) +
  stat_ecdf(alpha=1,geom="step", size = 0.75)+
  geom_vline(data=ddata.m,
             aes(xintercept = value,linetype=upper.lim),
             size=0.65) +
  geom_hline(yintercept = 0, linetype="solid", color = "black", size=0.5) +
  geom_hline(yintercept = 0.5, linetype="dashed", color = "black", size=0.65) +
  geom_hline(yintercept = 0.75, linetype="dashed", color = "black", size=0.65) +
  geom_hline(yintercept = 1, linetype="dashed", color = "black", size=0.65) +
  geom_vline(xintercept = 0, linetype="solid", color = "black", size=0.5) +
  geom_vline(xintercept = 50, linetype="dashed", color = "black", size=0.65) +
  geom_vline(xintercept = 75, linetype="dashed", color = "black", size=0.65) +
  scale_color_viridis(discrete = T, alpha=0.6, option="C",labels = tmp.lbs,direction = -1)+
  scale_fill_viridis(discrete = T, alpha=0.6, option="C",guide = 'none',direction = -1)+
  labs(color = 'Dataset',
       linetype = '',
       y = "Fn(x)", 
       x="|CBE|")+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(face="bold", size=14),
        axis.text.y = element_text(face="bold", size=14),
        axis.title.x = element_text(size=18),
        axis.title.y = element_text(size=18),
        legend.title = element_text(size=14),
        legend.text = element_text(size=11),
        panel.background = element_rect(fill='white', color='white',colour = 'white'))+
  scale_x_continuous(limits=c(0, 100), expand = c(0, 0), breaks=c(0,5,10,20,50,75,100))+
  scale_y_continuous(limits=c(0, 1), expand = c(0, 0))
)
```

